%% 05_wang_innovation.tex
\section{The Wang 2026 Innovation}
\label{sec:innovation}

\subsection{Overview}

The Wang~2026 paradigm adds exactly one new moving part to the
replicated model of Section~\ref{sec:agents}: a private subjective
valuation~$\hat{V}_i$, drawn from a ChatGPT call, that replaces the
utility agent's deterministic prior on the very first decision
tick. Everything downstream of the first tick---belief blending,
trust updates, utility maximisation, order matching, dividend
payments, and all of the market-quality instruments---is identical
to the pure Lopez-Lira run. The anchor is therefore a \emph{minimal}
perturbation of the decision model: the smallest change in
behaviour that still lets the population's psychology influence
its opening bids.

Two design choices make this minimal. First, the call is placed
once, in parallel across all utility agents, at run start, and the
result is consumed on the first decision tick and then cleared.
Second, the call never asks the language model for an \emph{order};
it asks only for a \emph{number}---a valuation per share in
experimental cents---which is routed through the same prior slot
used by equation~\eqref{eq:util_prior}. The agent's action is still
chosen by~\eqref{eq:eu}, using a wealth argument that the language
model never~sees. This construction isolates the psychological
channel from the trading channel; it also keeps the downstream
replay deterministic, because after the first tick no agent in
Wang~2026 depends on external~randomness of any kind.

\subsection{The anchor call}

Let $A_U$ be the set of utility agents in the current population.
For each $i\in A_U$ the simulator constructs a short prompt
$\pi_i$ that encodes three facts: the paper constants $(T,\E[d])$,
the agent's risk preference $U_i\in\{\Uloving,\Uneutral,\Uaverse\}$
and bias mode $b_i\in\{\text{over},\text{none},\text{under}\}$, and
the fact that the market is shared with DLM~2005 experienced
counterparties. The system prompt is fixed and instructs the model
to return a single number with no explanation, currency symbol, or
prose. Concretely, the user prompt template reads
\begin{quote}\small
Market: $T$ periods. Each share pays 0 or 20 cents with equal
probability per period, expected dividend $\E[d]$. Risk-neutral
fundamental value at run start is $\FV_0$ cents (declines by
$\E[d]$ per period). You, agent $i$, are \emph{risk} and
\emph{bias}. You share the market with Experienced agents
(Dufwenberg, Lindqvist and Moore 2005) who know the asset decays
to zero and refuse to overpay in late periods. What is your
private subjective valuation per share right now, in cents?
\end{quote}
where $\text{risk}\in\{\text{risk-loving},\text{risk-neutral},
\text{risk-averse}\}$ and $\text{bias}\in\{\text{persistently
optimistic},\text{unbiased},\text{persistently pessimistic}\}$. All
calls are fired in parallel as a single
$\texttt{Promise.all}$~batch, so the network latency is one round
trip rather than $|A_U|$ round trips.

The raw reply is parsed for the first signed decimal, parsed as a
float, and clamped to
\begin{equation}
\hat{V}_i\in[\,0.25\,\FV_0,\,1.75\,\FV_0\,]\;=\;[25,\,175],
\label{eq:clamp}
\end{equation}
so that any out-of-range reply is bounded and cannot destabilise
the downstream engine. A missing reply, a parse failure, or an HTTP
error silently drops the anchor for that agent; its prior then
defaults to equation~\eqref{eq:util_prior} on the first tick, and
the run is otherwise unaffected. This deterministic fallback
preserves the invariant that Wang~2026 produces a simulation with
or without network connectivity---a requirement for the design to be
reproducible and shareable.

\subsection{Consumption on the first decision tick}

On every utility agent's first call to \texttt{updateBelief} after
the simulation starts, the ``psych anchor'' slot is consulted in a
\emph{consume-and-clear} step:
\begin{equation}
V_i^{\text{prior}} \;=\; \begin{cases}
\hat{V}_i\cdot(1+\varepsilon_i) & \text{if $\hat V_i\ne\emptyset$,}\\[4pt]
\FV_t\cdot(1+b_i+\varepsilon_i) & \text{otherwise,}
\end{cases}
\qquad \text{then set }\hat V_i\leftarrow\emptyset.
\label{eq:consume}
\end{equation}
The first branch overrides the deterministic prior with the
language-model anchor and keeps the Lopez-Lira noise
$\varepsilon_i\sim\text{Uniform}[-0.03,0.03]$ so that two utility
agents assigned identical anchors still draw slightly different
first quotes. The second branch is the pure Lopez-Lira prior, used
on every tick after the first and on every tick when no anchor is
available. The ``clear'' step ensures that a second call to
\texttt{updateBelief} on the same tick, or the next decision tick
in the same run, goes through the standard path, so that the
language model's effect is felt at run~start and then decays
through the belief-blend equation~\eqref{eq:blend} as normal
messaging proceeds.

\subsection{Research question and instrumentation}

Let $i^\star=\arg\max_{i\in A_U}\hat{V}_i$ be the utility agent
with the highest anchored valuation, and define the top holder at
tick~$t$ as
\begin{equation}
j^\star(t)\;=\;\arg\max_{i\in A_U}\text{inventory}_i(t).
\label{eq:top}
\end{equation}
The research question is simply whether $j^\star(t)\to i^\star$ by
the end of the~run. The simulator reports, every tick, a boolean
\emph{match} indicator $\mathbb{1}\{j^\star(t)=i^\star\}$ and a
\emph{normalised gap}
\begin{equation}
g(t)\;=\;\frac{\hat{V}_{i^\star}-\hat{V}_{j^\star(t)}}{\hat{V}_{i^\star}}\;\in\;[0,1],
\label{eq:gap}
\end{equation}
together with the identities of the two agents and their
valuations. The gap collapses to zero when the asset has reached
the psychologically-maximal holder; it stays near~one when the
asset has migrated to an agent whose anchor is much lower than
the~maximum. Figure placeholders in the \texttt{figures/} directory
will host time-series plots of $g(t)$ and the match indicator for
each of the three paradigms.

The match indicator is related, but not identical, to the
allocative efficiency statistic
\begin{equation}
\AllocEff\;=\;\frac{\sum_i \hat{V}_i\cdot\text{inventory}_i(T)}
{\max_i\hat{V}_i\cdot Q},
\label{eq:ae}
\end{equation}
in which the numerator is the realised cross-agent utility weighted
by each agent's anchored valuation and the denominator is the
best-possible allocation in which every share is held by the
maximum-valuation agent. When $\AllocEff=1$ the asset has migrated to
$i^\star$ and the match indicator is true; but $\AllocEff$ can be close to
one even when the match indicator is false, for example when the
top holder is different from $i^\star$ but has a nearly-maximal
anchor. Reporting both statistics together is therefore
informative: the match indicator captures the identity of the
eventual owner, while $\AllocEff$ captures the continuous loss relative
to the psychological~optimum.

\subsection{Determinism, cost, and caching}

A single Wang~2026 run consumes $|A_U|$ chat completions as one
parallel batch and runs closed-form thereafter. With $|A_U|\le 6$
in the default preset, network cost is a handful of round trips on
the first frame and zero afterwards; no caching or retry layer is
needed because a missing anchor silently degrades to the
deterministic prior. The seeded PRNG makes identical anchor vectors
reproduce identical trajectories, so within-model replay is trivial
and across-model comparisons reduce to toggling the paradigm button.
