%% sections/abstract.tex
\begin{center}
\begin{minipage}{0.88\textwidth}
\small
\noindent\textbf{Abstract.} We replicate the finite-horizon experimental
asset market of \citet{dufwenberg2005bubbles} inside a deterministic
browser-based simulator and embed it with the expected-utility agent
of \citet{lopezlira2025llm}. Holding the market substrate, the agent
population, the prior, and the action choice rule fixed, we then
compare three distinct belief-update \emph{plans} for the utility
trader's per-period subjective valuation. \emph{Plan~I} is an
offline algorithmic blend of the agent's own prior with the mean
foreign peer claim, weighted by an endogenous experience counter
$k_i$ that ramps the weight from $0.60$ to $0.90$ across the
four-round session. \emph{Plan~II} replaces the blend with a
period-boundary chat-completion call whose prompt gives the language
model the closed-form risk-utility expression
$U\in\{\Uloving,\Uneutral,\Uaverse\}$ for that agent's risk type.
\emph{Plan~III} uses the same wiring as Plan~II but strips the
functional forms out of the prompt and only names the risk-preference
category. All three plans feed their output~$V_i^{\text{post}}$ into
the identical Lopez-Lira expected-utility scoring step, so any
difference in the resulting market dynamics is attributable to the
update channel alone. The simulator reports Haessel~$R^2$, normalised
deviation, amplitude, turnover, and allocative efficiency for all
three plans; the seeded mulberry32 PRNG makes within-plan replays
exact, and Plan~II and Plan~III silently degrade to Plan~I for any
agent whose LLM call fails, so every run can be reproduced without
network access. This paper documents the replication, the three
update plans, and the market-quality instruments used to compare
them.

\medskip
\noindent\textbf{Keywords:} experimental asset markets, bubbles,
expected utility, language models, belief updating, allocative
efficiency.

\smallskip
\noindent\textbf{JEL Codes:} C92, D84, G12, G41.
\end{minipage}
\end{center}
\bigskip
