%% sections/conclusion.tex
\section{Conclusion}
\label{sec:conclusion}

This paper documents a browser-based replication of the
experimental asset market of~\citet{dufwenberg2005bubbles},
extended with the expected-utility messaging population
of~\citet{lopezlira2025llm}, and introduces \emph{AI-Agent Prior
Elicitation} (AIPE), a protocol in which a single AI-agent call
seeds each utility trader's private subjective valuation at
run~start. The extension is deliberately minimal: one number per
agent, clamped to
$[0.25\,\FV_0,\,1.75\,\FV_0]$, written to the
\texttt{psychAnchor} slot and consumed on the first decision tick
of the standard Lopez-Lira prior. Everything downstream of the
first tick---the belief blend, the trust EMA, the expected-utility
action choice, the order matching, the dividend crediting, and
the market-quality instruments---is unchanged from the replicated
baseline, and the simulator's seeded mulberry32 generator ensures
that every run can be replayed identically with or without network
access.

The contribution is threefold. First, the testbed constitutes a
clean vanilla-JavaScript implementation of two otherwise unrelated
experimental designs, specified in enough detail
(Sections~\ref{sec:design}--\ref{sec:agents}) that a reader can
reconstruct the model from the paper alone. Second, AIPE
introduces a narrowly-scoped use of an AI agent as a
\emph{psychological prior elicitor} rather than as a trader: the
AI agent is never asked to choose an action, and its output is
routed through the same prior slot used by the deterministic
fallback, so that any behavioural effect is attributable to the
starting valuation rather than to an expansion of the action
space. Third, the research question---does the asset end up in
the hands of the agent with the highest anchored valuation---is
instrumented with a per-tick match indicator and a normalised
gap, and reported alongside the classical bubble statistics
(Haessel~$R^2$, normalised deviation, amplitude, turnover,
allocative efficiency). The instrumentation is designed so that
any replication of AIPE can report the same panel without
additional~code.

Three caveats shape the preliminary findings. First, the ranges in
Table~\ref{tab:mktq} are illustrative, not calibrated: a full
empirical test requires many replications per cell, a pooled
treatment of the network variance of the anchor, and a formal
baseline against a deterministic prior with matched mean and
variance. Second, the anchor is consumed exactly once, which bounds
the downstream signal at whatever the first-tick prior contributes
to the five-action expected-utility score; refreshing $\hat V_i$
each period would be a natural extension but is in tension with the
reproducibility invariant. Third, generalising beyond the DLM~2005
dividend distribution and horizon would require re-deriving the
market-quality instruments and re-calibrating the experienced
profiles of Table~\ref{tab:exp}.

The testbed is free of build steps and of third-party dependencies,
and ships with the architecture diagram, glossary, and slide deck
used in this paper. Its ``Experiment settings'' panel exposes every
numerical constant that shapes a run---the population mix, the
risk-preference share, the passive fill probability, the trust
learning rate, the bias amplitude, and the valuation noise---and
the seeded PRNG makes every edit to a constant observable as a
difference against a matched baseline. AIPE is the first extension
of this infrastructure that asks a question about AI agents, and
it opens several follow-ups: whether the anchor's effect on
allocative efficiency scales with the cross-agent variance
of~$\hat V$; whether different AI agents produce systematically
different anchor vectors for the same prompt; and whether adding
deceptive messaging amplifies or dampens the first-tick
stratification that the anchor produces.
