%% sections/design.tex
\section{Experimental Design}
\label{sec:design}

\subsection{Session structure: four consecutive rounds}

The unit of study in~\citet{dufwenberg2005bubbles} is a
\emph{session} of four consecutive ten-period markets played by the
same six traders. The paper insists on the terminology explicitly:
\begin{quote}
``A session involved four consecutive markets. In the following, we
shall talk in terms of four different rounds. Note the distinction
between rounds and periods; a round (being a market) consists of
ten periods.''~\citep[p.~1733]{dufwenberg2005bubbles}
\end{quote}
The session wrapper is the load-bearing element of the original
design: bubbles inflate in round~1, attenuate in rounds~2--3, and
have effectively died by round~4 even though the population is
identical---experience, not composition, kills the bubble. The
simulator therefore wraps the per-round market in a session loop of
$R=4$ rounds. At every round boundary the order book is cleared,
each agent's cash and inventory are rewound to its sampled spec,
and the period counter resets to one. Learned state---trust
matrices, belief modes, the EU candidate cache---is intentionally
preserved across the boundary so that the cross-round learning
channel that drives the paper's headline result is the only source
of persistent change between consecutive markets. Both
$R$ and $T$ are surfaced read-only in the simulator's ``Paper
constants'' panel and are not user tunable.

\subsection{The asset and the fundamental}

The asset is a finite-lived claim on an iid per-period dividend
that pays either zero or twenty experimental cents with equal
probability. Letting $d$ denote the period-$t$ dividend draw,
\begin{equation}
d\sim\text{Uniform}\{0,20\},\qquad \E[d]=10,\qquad T=10.
\label{eq:dividend}
\end{equation}
The asset lives for $T=10$ periods within each round and pays
nothing thereafter, so its risk-neutral fundamental value at the
start of period~$t$ of any round is
\begin{equation}
\FV_t \;=\; \E[d]\cdot(T-t+1) \;=\; 10\cdot(T-t+1),
\label{eq:fv}
\end{equation}
that is, a linearly declining staircase from $\FV_1=100$ to
$\FV_T=10$ with a terminal value of zero after the final dividend
pays. The fundamental resets to $\FV_1=100$ at the start of every
new round, so the price chart traces a four-tooth saw-tooth across
the full session. Equation~\eqref{eq:dividend} and
equation~\eqref{eq:fv} reproduce the dividend process and the
fundamental schedule of~\citet{dufwenberg2005bubbles} exactly. The
session length $R$, the per-round asset life $T$, and the dividend
mean $\E[d]$ are all surfaced read-only in the simulator's ``Paper
constants'' panel so that every comparison across populations holds
the market structure fixed.

\subsection{Continuous double auction with period discretisation}

The simulator discretises each period into $K=18$ decision rounds,
so a single round comprises $T\cdot K=180$ ticks and a full
four-round session comprises $R\cdot T\cdot K=720$ ticks. Every
tick, each agent posts at most one order, routed to a price--time
priority order book; a crossing pair matches at the resting price.
At the end of each period the outstanding quotes are withdrawn and
the dividend is drawn from~\eqref{eq:dividend} and credited to each
holder. At the end of period~$T$ of every non-final round the
session loop also rewinds endowments, clears the book, and resets
the period counter to one before the next round begins. The $K=18$
discretisation is the simulator's own approximation of the
two-minute continuous z-Tree window used in the paper, is pinned in
the ``Simulator constants'' panel, and is not user tunable.

\subsection{Strict-DLM mode: fixed $N=6$ and two discrete endowments}
\label{sec:strict-dlm}

The simulator's \emph{strict-DLM mode} replicates the original
protocol without relaxation. Selecting the DLM~2005 paradigm in the
navbar pins $N=6$, disables the population sliders, and routes the
sampling stage through a dedicated two-type draw. The paper
specifies exactly six traders per session:
\begin{quote}
``At each session, six subjects participated in a sequence of four
consecutive markets for an experimental asset.''~\citep[p.~1733]{dufwenberg2005bubbles}
\end{quote}
and exactly two discrete endowment types with identical implied
buy-and-hold value:
\begin{quote}
``Before a market opened, half of the traders started with 200 cents
and six assets, while each of the other traders started with 600
cents and two assets.''~\citep[p.~1733]{dufwenberg2005bubbles}
\end{quote}
Both bundles have buy-and-hold value $E[\text{cash}] = 200 +
6\cdot\E[d]\cdot T = 200 + 600 = 800$ for type~A and $600 +
2\cdot\E[d]\cdot T = 600 + 200 = 800$ for type~B, so the split is
strictly distributional: it varies the cash--inventory mix across
subjects without creating value heterogeneity.\footnote{More
directly, each bundle has the same risk-neutral fundamental value
$\mathrm{FV}_1 = 100$ per share at round start, so 200+6\,$\cdot$\,100 =
600+2\,$\cdot$\,100 = 800 + 200 = 1{,}000\,¢ once the initial cash is
counted.}

The simulator implements this with \texttt{dlmSampleAgents(rng)},
which pins six slots, assigns the index sequence
$[0,0,0,1,1,1]$ (three type~A, three type~B), Fisher--Yates
shuffles it against the dedicated sampling RNG, and writes the
matching cash/inventory pair onto each spec. No other mode touches
the two-type draw: Lopez-Lira and AIPE continue to use the
relaxed $U[800,1200]\times\{2,3,4\}$ sampler described for
backwards compatibility with their respective strategy-cube
populations.

\subsection{Endogenous experience}
\label{sec:endogenous-experience}

The original study is unusual in experimental economics in that
it treats experience as a \emph{procedural} rather than
compositional variable: subjects do not differ by type at the start
of a session, and no one is ever recruited because they are
``experienced''. Experience is a property of the session loop:
a subject acquires it by playing rounds~1, 2, and 3 alongside five
peers in a shared six-person market, and the paper's central
finding is that this procedural experience --- not a rule-change or
a population swap --- is what kills the bubble by round~4.

Reproducing DLM's design therefore requires that experience not
exist as a pre-assignable agent \emph{type}. The simulator uses a
single class, \texttt{DLMTrader}, with a two-state \texttt{decide}
method gated on an integer counter \texttt{roundsPlayed} that the
engine increments at every round boundary:
\begin{equation}
\mathtt{DLMTrader}.\texttt{decide}(\cdot) \;=\;
\begin{cases}
\texttt{inexperienced}(\cdot), & \text{if } \texttt{roundsPlayed} = 0, \\
\texttt{experienced}(\cdot),   & \text{if } \texttt{roundsPlayed} \ge 1.
\end{cases}
\label{eq:endo-exp}
\end{equation}
Every trader is instantiated with $\texttt{roundsPlayed}=0$, so
at the start of round~1 every participant is operating on the
inexperienced branch: it chases trend, accepts zero-intelligence
noise, ignores the fundamental horizon, and does not liquidate
late in a round. The engine's round-end hook then increments the
counter for every agent that has just finished a round, so the
same six traders enter round~2 having flipped to the experienced
branch, which anchors bids to $\varphi(h)\cdot\FV_t$ with horizon
$h=T-t+1$, refuses to buy above belief in late periods, and
liquidates its remaining inventory in the final three periods of a
round. Nothing else changes about the session; the same numeric
id, cash, inventory and trust state persist --- only the decision
branch flips.

Crucially, no agent is ever instantiated with a positive
\texttt{roundsPlayed}. The bubble-killing channel is therefore
purely procedural, matching~\citet{dufwenberg2005bubbles} exactly.

\subsection{Round-4 replacement and treatments}
\label{sec:round4}

The paper introduces two treatment conditions that manipulate the
mix of experienced and inexperienced traders at the start of
round~4:
\begin{quote}
``In the fourth round, depending on treatment, two or four
experienced subjects who had participated in the first three
rounds were randomly selected, removed, and replaced by the same
number of inexperienced subjects.''~\citep[p.~1733]{dufwenberg2005bubbles}
\end{quote}
The paper labels these two conditions by the \emph{fraction of
experienced subjects remaining} in round~4: \emph{R4-$\tfrac{2}{3}$}
for the condition in which four of the six traders are veterans
(two are replaced) and \emph{R4-$\tfrac{1}{3}$} for the condition
in which two of the six traders are veterans (four are
replaced).\footnote{See Table~1, footnote~a
\citep[p.~1734]{dufwenberg2005bubbles}---``Columns headed by 1--5
(6--10) correspond to sessions with two-thirds (one-third)
experienced traders in round~4''---and the hypothesis row of
Table~2 \citep[p.~1735]{dufwenberg2005bubbles}, which uses the
R4-$\tfrac{2}{3}$ / R4-$\tfrac{1}{3}$ shorthand in the
permutation and Wilcoxon--Mann--Whitney tests.}
The simulator fires the replacement between rounds~3 and~4
whenever strict-DLM mode is active. Let $k\in\{2,4\}$ denote the
number of \emph{fresh} traders spliced into round~4; we refer to
these two conditions as $\mathrm{T}2$ (paper's
R4-$\tfrac{2}{3}$, $k=2$) and $\mathrm{T}4$ (paper's
R4-$\tfrac{1}{3}$, $k=4$) in the rest of this paper, the UI, and
the source~code. The engine's round-end branch runs, in order:
\begin{enumerate}
\item snapshot each surviving agent's period-$T$ cash into
$\mathtt{roundFinalCash}[r-1]$ (Section~\ref{sec:payoff}),
\item log a \texttt{round\_end} event,
\item increment every surviving agent's \texttt{roundsPlayed} by
one, so the four or two survivors of the T2 (resp.~T4) sweep
carry \texttt{roundsPlayed}=3 into round~4,
\item Fisher--Yates select $k$ agents uniformly at random from the
pool of surviving non-replacement traders and remove them,
\item draw $k$ fresh \texttt{DLMTrader} specs from
\texttt{dlmSampleReplacementAgent} (each with a new name drawn
from the remaining name pool so that no two seats share a
display name, and a fresh endowment type~A or~B drawn iid) and
instantiate them with $\texttt{roundsPlayed}=0$ and a
\texttt{replacementFresh} flag for UI badging,
\item splice the new instances back into the engine at the
vacated numeric ids, so the order book, logger, and replay
system continue to key by a stable id,
\item rewind every surviving agent's cash and inventory to its
spec (the fresh agents take their own replacement endowment on
the same call), increment the round counter to 4, reset
\texttt{period=1}, and clear the book.
\end{enumerate}
The net effect is that round~4 is played with $6-k$ veterans on
the experienced branch of equation~\eqref{eq:endo-exp} and $k$
fresh subjects on the inexperienced branch --- precisely the
composition DLM's treatment manipulates.

\subsection{Multi-session batch}
\label{sec:multi-session}

\citet{dufwenberg2005bubbles} report ten sessions in total, five
per treatment, interleaved so that the overall design has
$10\times 4 = 40$ markets. The simulator provides a one-click
batch runner that reproduces this: pressing ``Run 10-session
batch'' in the DLM~panel drives ten synchronous
\texttt{Engine.runToEnd()} loops in two blocks of five,
alternating treatment, with an independent engine seed per
session and an independent two-type endowment draw. Each session
writes a compact summary row --- treatment, mean
$|p_t - \FV_t|/\FV_t$, per-share turnover, total trade count, mean
per-subject payoff, and the round-4 replacement record --- into a
results table in the DLM panel. The batch preserves the pre-batch
interactive state so the user's current session is not disturbed.

\subsection{Payoff accounting}
\label{sec:payoff}

Session payoffs in DLM~2005 have two components: the subject's
final cash at the end of each round, summed across all four
rounds, plus a show-up fee of \$5:
\begin{quote}
``Subjects were privately paid, in cash, the amount of their final
cash holdings from each round. They were also paid a show-up fee
of \$5.''~\citep[p.~1735]{dufwenberg2005bubbles}
\end{quote}
The simulator captures each surviving agent's cash at the end of
period~$T$ of every round via \texttt{Logger.logRoundFinalCash},
which writes a per-agent \texttt{\{id: cash\}} map into
$\mathtt{roundFinalCash}[r-1]$. For agent~$i$ the session payoff
is therefore
\begin{equation}
\pi_i \;=\; \sum_{r=1}^{R} \mathtt{roundFinalCash}[r-1][i] \;+\; 500,
\label{eq:payoff}
\end{equation}
where the $500$~cents show-up fee is added unconditionally and
shares held at the end of a round are worth nothing because the
asset's life has ended (the dividend stream has paid out). Agents
spliced in by the round-4 replacement step only contribute a
round-4 row to the sum because they did not play rounds~1--3; this
matches the paper's accounting for replacement subjects, who are
paid only for the single round they participate in plus the
show-up fee.

\subsection{Departures from the original protocol}
\label{sec:departures}

The simulator reproduces DLM~2005 without exception except in one
place: the original study runs a continuous two-minute open-outcry
double auction in z-Tree for each period~\citep[p.~1733]{dufwenberg2005bubbles},
whereas the simulator discretises that window into $K=18$ decision
rounds in order to run a deterministic tick-level loop that the
replay system can slice. Eighteen is dense enough to reproduce the
bubble--crash pattern at qualitative parity with the paper while
keeping the replay buffer compact; lower values starve the book,
higher values compress the per-tick update so agents barely
respond between decisions. This is the \emph{only} residual
simulator approximation. No other relaxation remains: the
population size, the two endowment types, the round count, the
dividend distribution, the fundamental schedule, the endogenous
experience gate, the round-4 treatment, the ten-session batch, and
the payoff + show-up fee structure are all in strict agreement
with the paper.

The Lopez-Lira and AIPE paradigms retain the $N\ge 4$ relaxed
sampler and the mixed F/T/R background described elsewhere in
this paper, since they are \emph{extensions} rather than
replications of DLM and therefore intentionally depart from the
strict design at the populations layer. Strict-DLM mode is the
single-button knob that turns every one of those relaxations
off.

\subsection{Reproducibility}

Every run is driven by a single scalar seed. A mulberry32 generator
$s_{k+1}=(s_k+\mathtt{0x6D2B79F5})\bmod 2^{32}$ produces the
$(0,1)$-uniform stream consumed by every stochastic step---dividend
realisations, zero-intelligence quotes, the noise term in the
utility prior, and the trust-update ordering. Calls to
\texttt{Math.random} from agent or market code are a contract
violation that breaks reproducibility by the (population, seed)
pair and are flagged as an invariant in the project style guide.
The seed is re-rolled only on Reset, and the replay system
reconstructs past states by slicing append-only history arrays on
\texttt{Market} and \texttt{Logger}---no frame is ever recomputed.
A single seed therefore reproduces the entire 720-tick session
end-to-end, rounds and round-end resets included.
