%% sections/model.tex
¥section{Model}
¥label{sec:model}

¥subsection{Asset and fundamental value}

The asset pays an iid per-period dividend
$d_t¥sim¥text{Uniform}¥{0,20¥}$ with $¥E[d_t]=10$ and lives for
$T=10$ periods per round.  The risk-neutral fundamental at the start
of period‾$t$ is
¥begin{equation}
¥FV_t = 10¥,(T-t+1),
¥label{eq:fv}
¥end{equation}
declining from $¥FV_1=100$ to $¥FV_T=10$.  Terminal value is zero.
The fundamental resets at every round boundary.  A session consists of
$R=4$ rounds; each period is discretised into $K=18$ ticks, giving
720 ticks per session.  The asset, dividend, and session structure
reproduce DLM exactly.

¥subsection{Market institution}

Trade takes place in a continuous double auction with price-time
priority.  Bids are sorted descending by price, asks ascending; ties
break by arrival time.  Each tick, every agent submits at most one
order.  A crossing pair matches at the resting price; self-matches are
prevented.  At the end of each period, all outstanding orders are
withdrawn and the dividend is drawn and credited.

¥subsection{Agent types}

Five agent types populate the market.  All observe the best bid‾$b_t$,
best ask‾$a_t$, last trade price‾$p_t$, and fundamental‾$¥FV_t$.

¥paragraph{Fundamentalist (F)}
Belief: $¥beta_t = ¥FV_t(1+0.04(r{-}0.5))$, a $¥pm2¥%$ perturbation.
Buys at‾$a_t$ when $a_t<0.98¥beta_t$; sells at‾$b_t$ when
$b_t>1.02¥beta_t$; otherwise posts passive quotes near $¥beta_t$ with
probability‾$0.45$ or holds.

¥paragraph{Trend follower (T)}
Estimates a slope from the last six in-round trade prices and
projects three ticks ahead.  Chases momentum when slope$¥,{>}¥,0$;
panic-sells when slope$¥,{<}¥,0$; liquidates excess inventory in
flat markets.  This is the positive-feedback channel
¥citep{delong1990positive}.

¥paragraph{Zero-intelligence (R)}
Posts $¥text{bid}¥sim¥text{Uniform}[0.7¥FV_t,1.1¥FV_t]$ and
$¥text{ask}¥sim¥text{Uniform}[0.9¥FV_t,1.3¥FV_t]$, following
¥citet{gode1993allocative}.  Ensures the book is never empty.

¥paragraph{DLM trader (D)}
Used only in strict-DLM mode ($N{=}6$).  A two-branch decision rule
is gated on the experience counter‾$k_i$, which starts at zero and is
incremented at each round boundary:
¥begin{equation}
¥text{decide}() =
¥begin{cases}
¥text{inexperienced rule}, & k_i = 0,¥¥
¥text{experienced rule}, & k_i ¥ge 1.
¥end{cases}
¥label{eq:dlm-gate}
¥end{equation}

¥emph{Inexperienced} ($k_i{=}0$).---The agent chases price momentum
with no fundamental anchor: when slope$¥,{>}¥,0$ it bids at
$p_t(1.02{+}0.04r)$; when slope$¥,{<}¥,0$ it asks at
$p_t(0.98{-}0.04r)$; otherwise it posts noisy quotes around‾$p_t$.
This branch generates the bubble.

¥emph{Experienced} ($k_i{¥ge}1$).---The agent anchors to a
horizon-discounted belief
$¥beta_t^{¥text{exp}} = ¥FV_t¥bigl(0.92 + 0.06¥,(T{-}t{+}1)/T¥bigr)$
and trades accordingly: it liquidates inventory when three or fewer
periods remain, refuses overpaying in late periods, crosses the book
only when the spread offers a discount or premium relative to
$¥beta_t^{¥text{exp}}$, and posts passive quotes near the belief with
probability‾$0.35$.  This branch kills the bubble.

¥paragraph{Utility agent (U)}
Follows ¥citet{lopezlira2025llm}.  On each tick the prior is
¥begin{equation}
V_i^{¥text{prior}} = ¥FV_t(1+b_i+¥varepsilon_{it}),
¥quad
b_i¥in¥{-0.15,¥,0,¥,0.15¥},
¥quad
¥varepsilon_{it}¥sim¥text{Uniform}[-0.03,0.03],
¥label{eq:util-prior}
¥end{equation}
where $b_i$ is a persistent optimistic, unbiased, or pessimistic bias.
The agent scores each action $a¥in¥{¥text{hold, cross bid, cross ask,
passive bid, passive ask}¥}$ by expected utility:
¥begin{equation}
¥EU(a) = p_f¥,U(w_1) + (1{-}p_f)¥,U(w_0),
¥qquad
a^¥star = ¥arg¥max_a¥EU(a),
¥label{eq:eu}
¥end{equation}
with constant fill probability $p_f{=}0.30$ and risk-preference
functionals normalised so $U(w_0){=}1$:
¥begin{equation}
¥Uloving(w) = (w/w_0)^2,¥qquad
¥Uneutral(w) = w/w_0,¥qquad
¥Uaverse(w) = ¥sqrt{w/w_0}.
¥label{eq:risk}
¥end{equation}

¥subsection{Communication and trust}

Each period, utility agents broadcast private valuations.  Three
deception modes govern the report: ¥emph{honest}
($¥hat v_i = v_i(1{¥pm}0.01)$), ¥emph{biased}
($¥hat v_i = v_i(1{¥pm}0.10)$, fixed sign), and ¥emph{deceptive}
($¥hat v_i = 0.82v_i$ when under-invested, $1.18v_i$ when
over-invested).  Pairwise trust $¥trust_{rs}¥in[0,1]$ is initialised
at‾$0.5$ and updated by exponential moving average against the
period's VWAP:
¥begin{equation}
¥trust_{rs} ¥leftarrow
(1{-}¥alpha)¥trust_{rs}
+ ¥alpha¥max¥!¥bigl(0,¥,1{-}|¥hat v_s{-}¥VWAP_t|/¥VWAP_t¥bigr),
¥qquad ¥alpha=0.30.
¥label{eq:trust}
¥end{equation}
Trust accumulates across rounds; all other per-round state resets.

¥subsection{Belief-update plans}

The period-boundary belief update maps the agent's prior and received
peer claims $¥msgs_i$ into a posterior $V_i^{¥text{post}}$ that
becomes next period's prior.  Three plans hold the upstream (prior,
messages, trust) and downstream (EU scoring, order matching, dividend
credit) stages fixed.

¥paragraph{Plan I (algorithmic blend)}
The posterior is the experience-weighted convex combination of
equation‾¥eqref{eq:plan1}: a novice ($k_i{=}0$) weights peers at
$40¥%$; a three-round veteran ($k_i{¥ge}3$) at‾$10¥%$.  If no
messages arrive, the prior is unchanged.  Plan‾I is deterministic
under the seeded PRNG.

¥paragraph{Plan II (LLM direct action with utility expression)}
At each period boundary, one language-model call per utility agent
bypasses equation‾¥eqref{eq:plan1} entirely.  The structured prompt
encodes market rules, the agent's status, decision-making principles,
role-specific guidance (risk-type and experience-level), and the
¥emph{closed-form} risk-utility expression:
$(w/w_0)^2$ for risk-loving, $w/w_0$ for neutral, $¥sqrt{w/w_0}$ for
averse.  The LLM returns a discrete action from the set
$¥{¥text{BUY¥_NOW}, ¥text{SELL¥_NOW}, ¥text{BID¥_1}, ¥text{BID¥_3},
¥text{ASK¥_1}, ¥text{ASK¥_3}, ¥text{HOLD}¥}$,
which the engine translates into an order using the current book state.
A failed or invalid action degrades silently to Plan‾I's EU evaluation.

¥paragraph{Plan III (LLM direct action with risk label only)}
Identical to Plan‾II except the functional form is deleted from the
prompt; only the category ¥emph{risk-loving}, ¥emph{risk-neutral}, or
¥emph{risk-averse} is named.  Every other input and the seven-action
output set are identical.

If Plans‾II and‾III yield different actions, the difference is
attributable to prompt structure alone.  If a run's LLM calls all
fail, the trajectory is identical to the matched Plan‾I run under the
same seed.
