%% sections/results.tex
¥section{Preliminary Results}
¥label{sec:results}

Because the simulator is designed for interactive replication
rather than a single canonical experiment, this section reports
preliminary findings from three paradigm presets rather than a
calibrated empirical test. The ¥emph{strict-DLM} paradigm runs the
exact protocol of‾¥citet{dufwenberg2005bubbles}: $N=6$ traders, two
discrete endowment types, four consecutive rounds, the round-4
replacement treatment, and the ten-session batch. The
¥emph{Lopez-Lira} and ¥emph{AIPE} paradigms continue to use the
relaxed population size and sampler, since they are extensions
rather than replications. Unless otherwise noted the seeded PRNG
is rolled fresh at the start of every run, and the AIPE anchor is
drawn from the default model with the paper's system and user
prompts verbatim. The market-quality statistics reported below are
pooled across all four rounds of a session; the per-round
breakdown that drives the experience-kills-bubbles finding is
described in the baseline DLM subsection.

¥subsection{Baseline strict-DLM: experience kills the bubble}

Under strict-DLM mode the six traders are all instantiated as
¥texttt{DLMTrader} instances with ¥texttt{roundsPlayed}=0, so every
subject in round‾1 operates on the inexperienced branch of
equation‾¥eqref{eq:endo-exp}. The round-1 tape therefore runs into
the bubble region: amplitude $A$ typically exceeds $0.35$, $¥ND$
exceeds $0.20$, and Haessel's $R^2$ is near zero because transaction
prices overshoot the staircase in periods‾2--5 and crash in
periods‾$8$--$10$. At the end of round‾1 the engine increments
every survivor's ¥texttt{roundsPlayed} to‾1 and the traders flip
to the experienced branch for round‾2. Amplitude falls, $¥ND$
contracts toward $0.10$, and $R^2$ rises to above $0.50$, even
though the population is identical, the seed is the same, and no
rule has changed. By round‾3 the bubble is essentially absent, and
round‾4 --- under either treatment --- matches the staircase to
within a few experimental cents for the veteran subset. The
simulator reproduces the qualitative DLM‾2005 finding without
pre-assigning experience at any point.

The round-4 replacement treatment manipulates the veteran/fresh
mix: T2 (the paper's R4-$¥tfrac{2}{3}$,
¥citealp[p.‾1735]{dufwenberg2005bubbles}) leaves four veterans and
two fresh subjects, T4 (the paper's R4-$¥tfrac{1}{3}$) leaves two
veterans and four fresh subjects. In preliminary runs, T4 widens
the round-4 price band compared to T2 because the four fresh
traders generate enough momentum-chasing on the inexperienced
branch to drag the book above the veterans' liquidation threshold
for a period or two before they dump; the size of the effect is
the research question the ten-session batch is built to measure.
Batch summaries print the mean normalised deviation and the mean
per-subject payoff (¥S¥ref{sec:payoff}, equation‾¥eqref{eq:payoff})
for each treatment so that side-by-side comparisons are possible
in a single click.

¥subsection{Lopez-Lira 2025: the strategy cube shapes the bubble}

Under the Lopez-Lira preset (six utility agents, no experienced),
the price dynamic depends sensitively on the population's
¥emph{strategy cube}: bias mode $¥times$ belief mode $¥times$ risk
preference. When $¥alpha_L>0.6$, the convex $¥Uloving(w)$ drives
aggressive ask-crossing in early periods with amplitude $A>0.40$,
positive residuals through periods $1$--$7$, and a terminal
collapse---the canonical SSW bubble-and-crash shape with
Haessel‾$R^2$ near zero. When $¥alpha_A>0.6$, concave $¥Uaverse(w)$
pulls residuals negative and raises $¥ND$ into the $0.15$--$0.25$
range, reflecting under-pricing rather than bubble formation.
Toggling the deceptive messaging mode produces the largest
perturbation: deceptive broadcasters drag the trust EMA down once
the period-end closeness update of‾¥eqref{eq:trust} runs, and the
mean-lie-magnitude statistic is tightly correlated with $¥ND$,
consistent with the closed-loop mechanism in
¥citet{lopezlira2025llm}.

¥subsection{AIPE: does the asset migrate to the highest anchor?}

The novel regime is the AIPE merged market. The Utility agents
sit alongside the fixed $F/T/R$ background; no agent is
pre-labeled as experienced. A live AI-agent anchor
vector‾$(¥hat V_i)_{i¥in A_U}$ is drawn at run start, and three
preliminary observations are clear from a handful of runs.

¥paragraph{Anchor dispersion.} The returned anchors are
heterogeneous across risk and bias archetypes in the expected
direction. Risk-loving, persistently-optimistic agents receive
anchors near the upper clamp‾$1.75¥,¥FV_0=175$; risk-averse,
persistently-pessimistic agents receive anchors near the lower
clamp‾$0.25¥,¥FV_0=25$; risk-neutral, unbiased agents cluster near
$¥FV_0=100$. The simulator reports the full anchor vector in its
subjective-valuation panel, so the reader can verify that the
model's psychology maps into‾$¥hat V$ in the direction the prompt
implies.

¥paragraph{First-tick quoting.} Because the anchor is consumed on
the first decision tick, the earliest orders in an AIPE run
are visibly stratified by anchor. Risk-loving agents with
$¥hat V¥approx175$ post aggressive bids near‾$¥FV_1=100$ and
cross the ask when possible; risk-averse agents with
$¥hat V¥approx 25$ post asks well below $¥FV_1$. The effect lasts
roughly one period before the Lopez-Lira belief-blending channel
pulls everyone back toward‾$¥FV_t$, but it produces a clear
opening asymmetry absent from a pure Lopez-Lira run with the same‾seed.

¥paragraph{The match indicator.} Across runs, the boolean
match indicator
$¥mathbb{1}¥{j^¥star(t)=i^¥star¥}$---whether the agent currently
holding the most shares is the agent with the highest anchor---is
not monotone in‾$t$. It typically starts near the seed endowment
distribution (which is independent of‾$¥hat V$), rises during
periods‾$1$--$3$ as the aggressive high-anchor agents accumulate
inventory, and then drifts in the terminal region as the
fundamentalist background liquidates its shares. In runs where
the highest-anchor agent is also the fundamentalist bubble's
counterparty, the indicator ends at one; in runs where the
background liquidates into an agent of middling anchor, the
indicator ends at zero and the
corresponding normalised gap‾$g(T)$ is typically in the
$0.10$--$0.30$ range. The allocative efficiency statistic $¥AllocEff$
moves in the same direction but stays strictly positive in every
run because the denominator in‾¥eqref{eq:ae} is‾$¥max_i¥hat
V_i¥cdot Q$ and the numerator always contains a nontrivial
contribution from the top holder.

¥subsection{Market-quality statistics at a glance}

Table‾¥ref{tab:mktq} summarises the canonical bubble statistics
under the three paradigms in qualitative---not yet calibrated---form.

¥begin{table}[h]
¥centering
¥small
¥caption{Market-quality statistics, qualitative ranges}
¥label{tab:mktq}
¥begin{tabular}{l c c c c}
¥toprule
& DLM r.¥,1 & DLM r.¥,4 & Lopez-Lira‾2025 & AIPE ¥¥
¥midrule
Haessel‾$R^2$       & $¥approx 0$  & $>0.80$      & $0.30$--$0.80$ & $0.40$--$0.85$ ¥¥
Normalised dev.‾$¥ND$& $>0.20$     & $<0.10$      & $0.10$--$0.30$ & $0.10$--$0.25$ ¥¥
Amplitude $A$       & $>0.35$      & $<0.15$      & $0.20$--$0.50$ & $0.20$--$0.45$ ¥¥
Turnover $¥TO$      & $0.6$--$1.1$ & $0.3$--$0.6$ & $0.6$--$1.2$   & $0.5$--$1.0$   ¥¥
Allocative eff.‾$¥AllocEff$& $0.6$--$0.8$& $¥approx 1$  & $0.6$--$0.9$   & $0.7$--$0.95$  ¥¥
Match $¥mathbb{1}¥{j^¥star(T)=i^¥star¥}$ & n/a & n/a & n/a & 50--70¥% runs ¥¥
¥bottomrule
¥end{tabular}
¥end{table}

The two strict-DLM columns ($r=1$ and $r=4$) together recover the
DLM finding: a large bubble in round‾1 with the same population
that produces a tight staircase in round‾4, separated only by the
three intervening rounds of procedural experience. The Lopez-Lira
and AIPE columns are pooled across the full session rather than
split by round because their relaxed populations do not have an
endogenous round-1 to round-4 sharpening channel --- the agents
keep the same trust/belief state across rounds but none of them
gate behaviour on ¥texttt{roundsPlayed}. AIPE sits between
strict-DLM round‾4 and Lopez-Lira because the fundamentalist
background still suppresses amplitude and improves‾$R^2$, while the
utility block introduces enough psychology to produce nontrivial
variation in the match indicator. Whether the match indicator
exceeds a naive baseline in a properly replicated experiment --- and
whether the anchor adds predictive power beyond the deterministic
prior --- is the question the testbed is built to answer.
